---
title: "5-report"
author: "Ben Bond-Lamberty"
date: "2/3/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("4-analysis.R")
```


Problems/thoughts:

* The Rh:Rs rise is too steep I think - implies a huge C cycle perturbation & v. high Q10
* Rs:MODIS and Rs:Beer GPP slopes are our best bet for a 'real' signal in this regard?
* But also Rh:GPPs, too
* If these are 'sensible' (???) then that strengthens things
* Look at this predicting for ...?


## Abstract

Here we show that:

* The _R_~H~:_R_~S~ ratio is rising over time, after controlling for climate, measurement method, and disturbance.
* The ratios of both _R_~H~ and _R_~S~ to remotely-sensed (MODIS) GPP are rising over time. Both trends are steeper for deciduous ecosystems.
* The ratios of both _R_~H~ and _R_~S~ to data synthesis (Beer et al.) GPP are rising over time. The _R_~S~ trend is steeper for deciduous ecosystems.
* The ratios of both _R_~H~ and _R_~S~ to SCIA SIF are rising over time. The _R_~S~ trend is steeper for deciduous ecosystems. No temporal trend is seen for the (very short) GOME2 SIF data, however.
* The ratio of _R_~S~ to FLUXNET GPP (in contrast to previous GPPs, a local measurement but smaller N) is rising in ecosystems that have experienced positive 1991-2010 precipitation trends.

Implications:

* This shows a consistent picture of rising _R_~H~, both in absolute terms and relative to _R_~S~ and GPP.
* This in turn implies a SOC feedback to the atmosphere driven by climate change.

To-do list:

* Check UMBS and Willow Creek soil respiration data - go to studies.
* The FLUXNET analysis is _so small N_. Plus, dominated by Harvard Forest.
* Look at SOC, given its importance (cf. Crowther 2016)
* Check bias due to partitioning method
* Read about regression with a ratio - problem?


SRDB status:

* `r nrow(srdb)` records used here, `r min(srdb$Study_midyear)` - `r max(srdb$Study_midyear)`
* `r sum(!is.na(srdb$Rs_annual))` annual _R_~S~ observations
* `r sum(!is.na(srdb$Rh_annual))` annual _R_~H~ observations

## Introduction

Perhaps the most urgent open question in terrestrial biogeochemistry and carbon cycling research, at least with respect to climate change, is the degree to which long-stored soil organic carbon (SOC) will be mineralized to the atmosphere.

Our [2010 Nature paper](http://www.nature.com/nature/journal/v464/n7288/full/nature08930.html) concluded by saying yeah, soil respiration (R~S~) is increasing globally, based on an analysis of a soil respiration database (SRDB), **but** we can't distinguish whether this is an acceleration of the C cycle, or a feedback:

>This positive global R~S~ response does not necessarily constitute a positive feedback loop between soil and the atmosphere; it could be driven by higher carbon inputs to soil rather than by mobilization of stored older carbon, as the temperature sensitivity and priming potential of recalcitrant carbon are uncertain.

Seven years later, we have a _lot_ more data, both within the SRDB and from other sources. How can we test whether heterotrophic respiration (R~H~) is in fact increasing?

R~H~ generally varies in a predictable way with R~S~ ([B-L 2004](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2486.2004.00816.x/abstract), [Subke 2006](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2486.2006.01117.x/abstract)). Everything else being equal, we would expect this ratio to change (increase) if soil C is being increasingly mineralized into R~H~. In addition, we would expect ecosystems' overall C balance to be affected (even though production might be itself increasing due to climate change).

## Methods Summary

We used an open-access soil respiration database (Bond-Lamberty and Thomson 2010), expanded and updated to include studies published through 2015. The database was filtered for studies that reported annual RH and/or RS; included explicit spatial and temporal data; took place in unmanaged or natural ecosystems without experimental manipulation; and used modern, standardized techniques to measure CO2. The result was a dataset (N = `r nrow(srdb)`, `r sum(!is.na(srdb$Rs_annual))` of _R_~S~ and `r sum(!is.na(srdb$Rh_annual))` of _R_~H~) of field-measured respiration fluxes between 1989 and 2014.


## Results

### SRDB analysis

The observed _R_~H~:_R_~S~ ratio is rising over time, from `r rh_rs_summary$rh_rs_mean[1]`±`r rh_rs_summary$rh_rs_sd[1]` in `r rh_rs_summary$yeargroup[1]` to `r rh_rs_summary$rh_rs_mean[nrow(rh_rs_summary)]`±`r rh_rs_summary$rh_rs_sd[nrow(rh_rs_summary)]` in `r rh_rs_summary$yeargroup[nrow(rh_rs_summary)]` (**Figure 1**).

Note for clarity this figure cuts off one extremely low-_R_~S~ point):

```{r, echo = FALSE}
print(p1_rh_rs)
```

This change is not significant using a Mann-Kendall trend test (`r pclean(mk1_rh_rs$sl)`, total N = `r nrow(s_rh_rs)`) but when ecosystem mean annual temperature (MAT) and mean annual precipitation (MAP) are controlled for in a linear model, a significant temporal trend exists (`r pclean(m1_rh_rs_signif)`; **Table 1**). This trend in the _R_~H~:_R_~S~ ratio was thus not induced by a site selection bias (as more productive sites typically exhibit higher _R_~H~:_R_~S~) over time. The regression also controls for researchers' sampling more disturbed sites, or sites with different leaf habit (deciduous/coniferous), or using different methods to partition _R_~H~ from _R_~S~ in recent years.

```{r, echo = FALSE}
knitr::kable(anova(m1_rh_rs), digits = 3)
```

This makes sense.

This is in a sense unsurprising, as from basic biokinetics (ref) we expect that the absolute value of observed annual _R_~H~ should respond to climate changes over the `r max(srdb$Study_midyear) - min(srdb$Study_midyear) + 1`-year record. This is true: in a linear regression mean annual precipitation (`r pclean(m2_rh_climate_map_signif)`) was the strongest driver, with potential evapotranspiration, stage (i.e. disturbance effect) and temperature anomaly (departure from the 1961-1990 mean) highly significant (all P < 0.001) as well.


### Remote sensing analyses

A rising _R_~H~:_R_~S~ ratio could be due to SOC loss and thus a climate feedback, or conceivably simply an acceleration of the C cycle (B-L 2010), if increased gross primary production (GPP) were for some reason preferentially emerging as _R_~H~. To test for this difference, we examined the ratio of soil-derived respiration to GPP, the ultimate source for both the autotrophic and heterotrophic soil surface CO~2~ fluxes. If SOC losses are increasing _R_~H~, one would expect both _R_~H~ and the total surface flux _R_~S~ to rise _faster_ than GPP. This is in fact the case (**Figure 2**).

```{r, echo = FALSE}
print(p_gppsif)
```

(Note that for clarity, this graph does not show a number of high RS:GPP and RS:SIF points. It also excludes (removes from the data set) `r nrow(s_gppsif_excluded)` points with R:GPP > `r MAX_FLUX_TO_GPP`.)

We perform a rigorous Mann-Kendall trend test, that does not account for climate, and fit a linear model with leaf habit, partition method, mean annual temperature, mean annual precipitation, and time (the trend test) as interacting factors:

```{r, echo = FALSE}
knitr::kable(rs_results)
```

Both _R_~H~ (the variable of primary interest, but for which many fewer measurements are available) and _R_~S~ (which incorporates both _R_~H~ and belowground autotrophic respiration) exhibit significant trends over time (`r pclean(m_Beer.GPP_Rs_annual_signif)` for _R_~S~:Beer GPP, P < 0.001 for all others) relative to two different GPP data sources. The _R_~S~:MODIS GPP ratio may be rising faster in deciduous ecosystems than in evergreen ecosystems (`r pclean( m_MODIS.GPP_Rs_annual_leaf_trend_signif)`).

We also examined the ratios of _R_~H~ and _R_~S~ to two sun-induced fluorescence (SIF) global data products, which are correlated with GPP but provide a completely independent test. Both _R_~H~:SIF and _R_~S~:SIF exhibited temporal trends against the SCIAMACHY (http://www.sciamachy.org) data after controlling for disturbance and measurement method (P < 0.001 for both; **Figure 2**). The _R_~H~ rise was significantly stronger in deciduous ecosystems (`r pclean(m_SCIAMACHY.SIF_Rh_annual_leaf_trend_signif)`). No _R_~H~:SIF or _R_~S~:SIF temporal trend was found using the GOME2 SIF data, but this is by far the shortest and most recent (2007-2014) of the datasets tested.


### FLUXNET analysis

One potential objection to the above analysis is that there's an enormous spatial mismatch between remotely-sensed data and the chamber measurements used to compute _R_~H~ and _R_~S~. To address this problem, we use FLUXNET Tier 1 data to identify GPP measurements made in the same location and year as _R_~S~ measurements. The advantage is this is a local (<= `r MAX_FLUXNET_DIST` km), empirical, and highly standardized GPP measurement. The tradeoff, as compared to a global GPP product, is that it drastically reduces the number of available data points; a FLUXNET-_R_~H~ comparison is not possible.

```{r, echo = FALSE}
print(p_fluxnet)
```

The M-K overall trend is significant (Mann-Kendall `r pclean(mk3_fluxnet$sl)`).

```{r, echo = FALSE}
print(anova(m_fluxnet))
hf_percnt <- round(nrow(s_fluxnet_nohf) / nrow(s_fluxnet) * 100, 0)
```

One potential problem is that Harvard Forest has an enormous number of points - `r hf_percnt`% - but it exhibits a falling _R_~S~:GPP ratio over time. Excluding HF gives:

```{r, echo = FALSE}
print(anova(m_fluxnet_nohf))
```

In both cases, there's a significant temporal trend in the _R_~S~ to GPP ratio after controlling for stage (aggrading/mature), leaf habit (deciduous/evergreen), mean annual temperature, and mean annual precipitation.

## Summary

Test  | N | Leaf habit? | Precipitation?		
------| - | ------------------- | -------------		
↑ R~H~:R~S~ ratio over time  | `r nrow(s_rh_rs)` | ?	| Yes	
↑ R~H~:Beer GPP over time  | `r nrow(s_Beer.GPP_Rh_annual)` | Yes	| Yes
↑ R~S~:Beer GPP over time  | `r nrow(s_Beer.GPP_Rs_annual)` | Yes | Yes		
↑ R~H~:MODIS GPP over time  | `r nrow(s_MODIS.GPP_Rh_annual)` | Yes	| Yes
↑ R~S~:MODIS GPP over time  | `r nrow(s_MODIS.GPP_Rs_annual)` | Yes | Yes
↑ R~H~:SCIA SIF over time  | `r nrow(s_SCIAMACHY.SIF_Rh_annual)` | Yes	| Yes
↑ R~S~:SCIA SIF over time  | `r nrow(s_SCIAMACHY.SIF_Rs_annual)` | Yes | Yes		
↑ R~H~:GOME2 SIF over time  | `r nrow(s_GOME2.SIF_Rh_annual)` | Yes	| Yes
↑ R~S~:GOME2 SIF over time  | `r nrow(s_GOME2.SIF_Rs_annual)` | Yes | Yes
↑ R~S~:FLUXNET GPP over time  | `r nrow(s_fluxnet)` | Yes | Yes		
R~H~ climate response | `r length(m2_rh_climate$residuals)` | ? | Yes

### Potential source of error

**Site selection bias.** Researchers might be sampling more recently-disturbed sites over time; we control this by testing for the interaction of `Year` and `Stage`, a field in the SRDB that records "Aggrading" or "Mature". Another possibility is preferential sampling of more-productive sites over time, with higher _R_~H~:_R_~S~, and we control for this by testing the _R_~H~:_R_~S~ trend against MAT and MAP.

**Measurement bias.** We use only IRGA has gas chromatography methods from 1989 onwards. These are standardized and modern techniques.

**Partitioning method bias.** We test for this effect.

**Autocorrelation**. FLUXNET GPP uses _R_~S~ at least sometimes? And that feeds into Beer et al. upscaling product. Too bad; GPP, in particular FLUXNET GPP, is the gold standard we have to compare against.

**Site correlation**. The FLUXNET analysis should probably include site as a random effect to control for this. TODO.

**Random error.** We always test for influential outliers. But this is a correlative, observational study, and the possibility of a Type I error can never be completely excluded. There's a lot of variability.
