---
title: "5-report"
author: "Ben Bond-Lamberty"
date: "2/3/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("4-analysis.R")
```

## Abstract

Here we show that:

* The _R_~H~:_R_~S~ ratio is rising over time, after controlling for climate, measurement method, and disturbance.
* The ratios of both _R_~H~ and _R_~S~ to remotely-sensed (MODIS) GPP are rising over time. Both trends are steeper for deciduous ecosystems.
* The ratios of both _R_~H~ and _R_~S~ to data synthesis (Beer et al.) GPP are rising over time. The _R_~S~ trend is steeper for deciduous ecosystems.
* The ratio of _R_~S~ to FLUXNET GPP (in contrast to previous GPPs, a local measurement but smaller N) is rising in ecosystems that have experienced positive 1991-2010 precipitation trends.

Implications:

* This shows a consistent picture of rising _R_~H~, both in absolute terms and relative to _R_~S~ and GPP.
* This in turn implies a SOC feedback to the atmosphere driven by climate change.

To-do list:

* Update SRDB through 2015
* Incorporate a SIF analysis
* FLUXNET matching should depend on DBF, ENF, etc.
* Look at SOC, given its importance (cf. Crowther 2016)
* Check bias due to partitioning method
* Read about regression with a ratio - problem?
* Make better use of anova()

SRDB status:

* `r nrow(srdb)` records used here, `r min(srdb$Study_midyear)` - `r max(srdb$Study_midyear)`
* `r sum(!is.na(srdb$Rs_annual))` annual _R_~S~ observations
* `r sum(!is.na(srdb$Rh_annual))` annual _R_~H~ observations

## Introduction

Perhaps the most urgent open question in terrestrial biogeochemistry and carbon cycling research, at least with respect to climate change, is the degree to which long-stored soil organic carbon (SOC) will be mineralized to the atmosphere.

Our [2010 Nature paper](http://www.nature.com/nature/journal/v464/n7288/full/nature08930.html) concluded by saying yeah, soil respiration (R~S~) is increasing globally, based on an analysis of a soil respiration database (SRDB), **but** we can't distinguish whether this is an acceleration of the C cycle, or a feedback:

>This positive global R~S~ response does not necessarily constitute a positive feedback loop between soil and the atmosphere; it could be driven by higher carbon inputs to soil rather than by mobilization of stored older carbon, as the temperature sensitivity and priming potential of recalcitrant carbon are uncertain.

Seven years later, we have a _lot_ more data, both within the SRDB and from other sources. How can we test whether heterotrophic respiration (R~H~) is in fact increasing?

R~H~ generally varies in a predictable way with R~S~ ([B-L 2004](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2486.2004.00816.x/abstract), [Subke 2006](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2486.2006.01117.x/abstract)). Everything else being equal, we would expect this ratio to change (increase) if soil C is being increasingly mineralized into R~H~. In addition, we would expect ecosystems' overall C balance to be affected (even though production might be itself increasing due to climate change).

## Methods and results

### SRDB analysis

The _R_~H~:_R_~S~ ratio is rising over time in the SRDB (total N = `r nrow(s_rh_rs)`; note for clarity this figure cuts off one extremely low-_R_~S~ point):

```{r, echo = FALSE}
print(p1_rh_rs)
```

This rise is not significant using a basic Mann-Kendall trend test (`r pclean(mk1_rh_rs$sl)`) but is once we control for climate:

```{r, echo = FALSE}
print(summary(m1_rh_rs))
```

This says that a temopral trend exists (`r pclean(m1_rh_rs_trend)`) independent of sites' MAT and MAP, i.e. this trend was not induced by a site selection bias (e.g. more productive sites that typically exhibit higher _R_~H~:_R_~S~) over time. The regression also tests for researchers sampling more disturbed sites, or sites with different leaf habit (deciduous/coniferous), in recent years; this factor is not significant.

### Remote sensing analyses

Perhaps _R_~H~:_R_~S~ is rising over time, but this could be due to either an acceleration of the C cycle, or SOC loss and thus a climate feedback (cf. [Bond-Lamberty and Thomson 2010](http://www.nature.com/nature/journal/v464/n7288/full/nature08930.html)). One way to test for this difference would be to look at the ratio of soil respiration (either _R_~H~ and _R_~S~) to gross primary production (GPP), the ultimate source for both the autotrophic and heterotrophic soil surface CO~2~ fluxes. If increased SOC losses are feeding the soil respiration fluxes, we would expect them to rise _faster_ than GPP.

```{r, echo = FALSE}
print(p_gpp_remotesensing)
```

(Note that for clarity, this graph does not show a number of high RS:GPP points. It also excludes (removes from the data set) `r nrow(s_gpp_excluded)` points with R:GPP > `r MAX_FLUX_TO_GPP`.)

We both perform a rigorous Mann-Kendall trend test, that does not account for climate, and fit a linear model with leaf habit, mean annual temperature, mean annual precipitation, and time (the trend test) as interacting factors. Generally:

Resp. flux | GPP source  | M-K trend? | lm trend? | Higher deciduous trend?
---------------- | ----------- | ---------- | --------- | -----------------
_R_~S~ | MODIS       | `r pclean(mk_gpp_modis_rs$sl)`   | Yes    | Yes
_R_~H~ | MODIS       | `r pclean(mk_gpp_modis_rh$sl)`   | Yes    | Yes
_R_~S~ | Beer et al. | `r pclean(mk_gpp_beer_rs$sl)`   | Yes    | Yes
_R_~H~ | Beer et al. | `r pclean(mk_gpp_beer_rh$sl)`   | Yes    | No

Both _R_~H~ (the variable of primary interest, but many fewer measurements) and _R_~S~ (which incorporates both _R_~H~ and belowground _R_~A~) seem to be rising relative to GPP with time, and doing so faster in deciduous ecosystems.

#### SIF

To do.

### FLUXNET analysis

A reasonable objection to the above is that there's an enormous spatial mismatch between remotely-sensed data and the chamber measurements used to compute _R_~H~ and _R_~S~. To address this problem, we use FLUXNET Tier 1 data to identify GPP measurements made in the same location and year as _R_~S~ measurements. The advantage is this is a local (<= `r MAX_FLUXNET_DIST` km), empirical, and highly standardized GPP measurement. The tradeoff, as compared to a global GPP product, is that it drastically reduces the number of available data points; a FLUXNET-_R_~H~ comparison is not possible.

```{r, echo = FALSE}
print(p_fluxnet)
```

Almost all FLUXNET sites experienced 1991-2010 warming trends, but  precipitation trends were more variable, and the graph above is segmented by whether the site became drier (left) or wetter (right). Grey trend lines correspond to repeated measurements at individual FLUXNET sites. The M-K overall trend for sites with precipitation increases, on the right, is not significant (Mann-Kendall `r pclean(mk3_fluxnet_wet$sl)`), but is in a linear model once we control for climate and, in particular, stage:

```{r, echo = FALSE}
print(summary(m_fluxnet_wet))
hf_percnt <- round(length(m_fluxnet_wet_nohf$residuals) / length(m_fluxnet_wet$residuals) * 100, 0)
```

One potential problem is that Harvard Forest has a enormous number of points - `r hf_percnt`% of the 'wetter' sites, and it exhibits a falling _R_~S~:GPP ratio over time. Excluding HF gives:

```{r, echo = FALSE}
print(summary(m_fluxnet_wet_nohf))
```

In this case, the Mann-Kendall trend test is significant (`r pclean(mk3_fluxnet_wet_nohf$sl)`). In both cases, there's a significant temporal trend in the _R_~S~ to GPP ratio after controlling for stage (aggrading/mature), leaf habit (deciduous/evergreen), mean annual temperature, and mean annual precipitation.

## Summary

Test  | N | Deciduous stronger? | Consistent with R~H~ feedback?		
------| - | ------------------- | -------------		
↑ R~H~:R~S~ ratio over time  | `r nrow(s_rh_rs)` | ?	| Yes	
↑ R~H~:Beer GPP over time  | `r nrow(s_gpp_beer_rh)` | Yes	| Yes
↑ R~S~:Beer GPP over time  | `r nrow(s_gpp_beer_rs)` | Yes | Yes		
↑ R~H~:MODIS GPP over time  | `r nrow(s_gpp_modis_rh)` | Yes	| Yes
↑ R~S~:MODIS GPP over time  | `r nrow(s_gpp_modis_rs)` | Yes | Yes
↑ R~S~:SIF over time  | ? | ? | ?
↑ R~S~:FLUXNET GPP over time  | `r nrow(s_fluxnet_wet)` | Yes | Yes		
R~H~ climate response | `r length(m2_rh_climate$residuals)` | ? | Yes

### Potential source of error

**Site selection bias.** Researchers might be sampling more recently-disturbed sites over time; we control this by testing for the interaction of `Year` and `Stage`, a field in the SRDB that records "Aggrading" or "Mature". Another possibility is preferential sampling of more-productive sites over time, with higher _R_~H~:_R_~S~, and we control for this by testing the _R_~H~:_R_~S~ trend against MAT and MAP.

**Measurement bias.** We use only IRGA has gas chromatography methods from 1989 onwards. These are standardized and modern techniques.

**Partitioning method bias.** TODO - need to check this more rigorously.

**Autocorrelation**. FLUXNET GPP uses _R_~S~ at least sometimes? And that feeds into Beer et al. upscaling product. Too bad; GPP, in particular FLUXNET GPP, is the gold standard we have to compare against.

**Site correlation**. The FLUXNET analysis should probably include site as a random effect to control for this. TODO.

**Random error.** We always test for influential outliers. But this is a correlative, observational study, and the possibility of a Type I error can never be completely excluded. There's a lot of variability.
